---
title: "Need for Cognition Example Dataset"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{NeedforCognition}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The following vignette showcases the use of `elfs()` when applied to data from the Need for Cognition scale (Cacioppo & Petty, 1982), using data collected and originally analyzed by Hussey and Hughes (2020). First, the dataset is loaded.

```{r setup, message = FALSE}
library(elfs)
library(dplyr)
library(magrittr)

nfc_data <- elfs::NeedforCognition_HusseyHughes_data
```

## One-factor Model

For our first model, we conduct the most simplistic model on the scale (all 18 items load onto the same factor).

```{r, message = FALSE}
nfc_elf <- elfs(data = nfc_data,
                f1_cols = starts_with("nfc"), f1_name = "NeedforCognition",
                chrome_bypass = T)
```

The resulting `elfs()` output contains information about your fitted model. By initiating specific calls, you can output information regarding:

```{r}
#Model Fit
nfc_elf$fit

#Individual Factor Reliabilities
nfc_elf$rel

#Latent Factor Correlations
nfc_elf$corr

#Comparison to Tau-equivalent Model Assumed by Cronbach's Alpha
nfc_elf$tau
```

If latent factor scores are ultimately decided on, they can be easily extracted and incorporated into a dataset (here, the original dataframe):

```{r}
head(nfc_elf$lfs, 10)

nfc_clean <- nfc_data %>%
  bind_cols(nfc_elf$lfs)
glimpse(nfc_clean)
```

## Measurement Invariance

This simple model can be expanded, firstly by testing for measurement invariance. Within `nfc_data`, "sex" is a binary variable denoting the sex of the participant. Separate models for each group can be run using the following:

```{r, message = FALSE}
nfc_invar <- elfs(data = nfc_data,
                  f1_cols = starts_with("nfc"), f1_name = "NeedforCognition",
                  meas_invar = "sex",
                  chrome_bypass = T)
```

The resulting information regarding measurement invariance can be viewed:

```{r}
#Raw model
nfc_invar$group_models

#Table summary
nfc_invar$meas_invar
```

If desired, individual latent factor scores can be extracted for each group. 

```{r}
#Scores for sex = "m"
head(nfc_invar$group_lfs[[1]]$m, 10)

#Scores for sex = "f"
head(nfc_invar$group_lfs[[1]]$f, 10)
```

## Dynamic Fit Indices (McNeish & Wolf, 2023)

Finally, dynamic fit indices can also be requested:

```{r, message = FALSE}
nfc_dynamic <- elfs(data = nfc_data,
                    f1_cols = starts_with("nfc"), f1_name = "NeedforCognition",
                    dynamic = T,
                    chrome_bypass = T)
```

Resultant information about updated model cutoffs can be displayed:

```{r}
nfc_dynamic$dynamic
```
