---
title: "Short Dark Triad Example Dataset"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{ShortDarkTriad}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The following vignette showcases the use of `elfs()` when applied to data from the Short Dark Triad (SD3; Paulhus & Jones, 2001), collected from the Open-Source Psychometrics Project. First, the dataset is loaded; responses were filtered to retain complete data, after converting responses of 0 to NA.

```{r setup, message = FALSE}
library(elfs)
library(dplyr)
library(magrittr)

SD3_data <- elfs::ShortDarkTriad_data
```

## Multi-factor Model

For our first model, we can conduct the most simplistic model on the SD3. Treating each of the factors of the Dark Triad (Machiavellianism, Narcissism, and Psychopathy) as separate factors measured by 9 items each. In the following model, all latent factors are allowed to be covary (i.e., oblique rotation).

```{r, message = FALSE}
#column selection can take in a traditional character vector, or use "starts_with()" or "ends_with()"
SD3_elf_oblique <- elfs(data = SD3_data,
                        f1_cols = starts_with("M"), f1_name = "Machiavellianism",
                        f2_cols = c("N1", "N2", "N3", "N4", "N5", "N6", "N7", "N8", "N9"), f2_name = "Narcissism",
                        f3_cols = starts_with("P"), f3_name = "Psychopathy", 
                        chrome_bypass = T)
```

The resulting `elfs()` output contains information about your fitted model. By initiating specific calls, you can output information regarding:

```{r}
#Model Fit
SD3_elf_oblique$fit

#Individual Factor Reliabilities
SD3_elf_oblique$rel

#Latent Factor Correlations
SD3_elf_oblique$corr

#Comparison to Tau-equivalent Model Assumed by Cronbach's Alpha
SD3_elf_oblique$tau
```

If latent factor scores are ultimately decided on, they can be easily extracted and incorporated into a dataset (here, the original dataframe):

```{r}
head(SD3_elf_oblique$lfs, 10)

SD3_clean <- SD3_data %>%
  bind_cols(SD3_elf_oblique$lfs)
glimpse(SD3_clean)
```

## Ordinal Measured Variables

An additional specification that measured variables should be treated as ordinal variables (i.e., using the DWLS estimator, instead of ML) can be specified using the typical `ordered = T` syntax from *lavaan*.

```{r, message = FALSE}
#column selection can take in a traditional character vector, or use "starts_with()" or "ends_with()"
SD3_elf_ordered <- elfs(data = SD3_data,
                        f1_cols = starts_with("M"), f1_name = "Machiavellianism",
                        f2_cols = starts_with("N"), f2_name = "Narcissism",
                        f3_cols = starts_with("P"), f3_name = "Psychopathy", 
                        ordered = T, 
                        chrome_bypass = T)
```

Note that when `ordered = T`, nonconvergence in the tau-equivalent model is much more likely, restricting potential comparisons.

The resulting tables displaying model fit are modified accordingly:

```{r}
#Model Fit
SD3_elf_ordered$fit

#Comparison to Tau-equivalent Model Assumed by Cronbach's Alpha
SD3_elf_ordered$tau
```

## Bi-factor Model

`elfs()` also contains baseline functionality to fit bifactor models. For example, we can specify a general "Evil" factor explaining leftover variance in the original model:

```{r, message = FALSE}
SD3_elf_bifac <- elfs(data = SD3_data,
                      f1_cols = starts_with("M"), f1_name = "Machiavellianism",
                      f2_cols = starts_with("N"), f2_name = "Narcissism",
                      f3_cols = starts_with("P"), f3_name = "Psychopathy", 
                      fg_cols = c(starts_with("M"), starts_with("N"), starts_with("P")), fg_name = "Evil",
                      chrome_bypass = T)
```

The resulting models must set covariances between latent factors to 0 to identify the model. The `elfs()` output is then updated:

```{r}
#Model Fit
SD3_elf_bifac$fit

#Individual Factor Reliabilities
SD3_elf_bifac$rel

#Latent Factor Correlations
SD3_elf_bifac$corr

#Comparison to Tau-equivalent Model Assumed by Cronbach's Alpha
SD3_elf_bifac$tau
```

As before, if latent factor scores are decided on, they can be easily extracted and attached to any dataset:

If latent factor scores are ultimately decided on, they can be easily extracted and incorporated into a dataset (here, the original dataframe):

```{r}
head(SD3_elf_bifac$lfs, 10)

SD3_clean_bifac <- SD3_data %>%
  bind_cols(SD3_elf_bifac$lfs)
glimpse(SD3_clean_bifac)
```
